## Peer Assessment 1

=========================

###Loading the data

Unzip and then load the data from working directory.
```{r LoadData, echo = TRUE}
require(lattice)
data0 <- read.csv("C:/Documents and Settings/5241753/Desktop/Kazi/Coursera/Reproducible Research/Assessment 1/activity.csv")
```


###What is mean total number of steps taken per day?

Calculate the total number of steps pers day.
```{r Aggregation, echo = TRUE}
StepsPerDay <- aggregate(data0$steps, by = list(data0$date), "sum")
```

Plotting a histogram of the total steps per day.
```{r Histogram, echo = TRUE}
hist(StepsPerDay[,2], breaks = 30, xlab = "Steps",
     main = "Histogram of Total Number of Steps Taken per Day")
```

Calculate the mean and median of the total number of steps per day.
```{r Mean and Median, echo = TRUE}
StepMean    <-  as.integer(mean(StepsPerDay[,2], na.rm = TRUE))
StepMedian  <-           median(StepsPerDay[,2], na.rm = TRUE)
```

The mean and median total number of steps taken per day are: `r StepMean` and `r StepMedian` respectively. 


###What is the average daily activity pattern?

Create a time series plot from the data provided.
```{r Time Series Plot, echo = TRUE}

AveStepsPerInt <- aggregate(data0$steps[!is.na(data0$steps)],
                            by = list(data0$interval[!is.na(data0$steps)]), mean)

plot(unique(data0$interval), AveStepsPerInt[,2], type = 'l', xlab = '5 Min Intervals', ylab = 'Ave Steps per Interval')
```

calculate which 5-minute interval contains the most steps, on average. 
```{r Max Steps, echo = TRUE}
MaxStep <- AveStepsPerInt[ AveStepsPerInt[,2] == max(AveStepsPerInt[,2]),1]
```

The 5-min interval that contains (on average) the most steps is the `r MaxStep` interval.


###Imputing missing values

Count the total number of rows that have NAs
```{r Missing Values, echo = TRUE}
table(is.na(data0$steps))[[2]]
```

Filling in the blanks
```{r Filling in NAs, echo = TRUE}
data1 <- data0[,]
data1 <- merge(x = data0, y = AveStepsPerInt, by.x = "interval", by.y = "Group.1", all = TRUE)
data1 <- data1[order(data1$date, data1$interval),]
data2 <- data1[,c(3,1,2,4)]
data2 <- unique(data2)
row.names(data2) <- NULL 
data2[is.na(data2$steps) ,3] <- round(as.numeric(data2[is.na(data2$steps) ,4]),0)
data2 <- data2[,-4]

````


###New results 

Creating a histogram, calculating mean and median based of the new dataset
```{r New histogram and mean/median, echo =TRUE}
StepsPerDay <- aggregate(data2$steps, by = list(data2$date), "sum")

hist(StepsPerDay[,2], breaks = 30, xlab = "Steps",
     main = "Histogram of Total Number of Steps Taken per Day")

StepMean    <-  as.integer(  mean(StepsPerDay[,2], na.rm = TRUE))
StepMedian  <-  as.integer(median(StepsPerDay[,2], na.rm = TRUE))
```
The new mean and new median total number of steps taken per day are: `r StepMean` and `r StepMedian` respectively. 

###Weekday and Weekend patterns
```{r Weekday & Weekends, echo = TRUE}
data3     <- data2[,]
data3[,4] <- weekdays(as.Date(data2[,1]))
data3[,5] <- "Weekday"
data3[which(data3$V4 == "Saturday" | data3$V4 == "Sunday"),5] <- "Weekend"

AveStepsPerDayType <- aggregate(data3$steps,by = list(data3$V5,data3$interval), mean)

xyplot(x~Group.2 | Group.1, data=AveStepsPerDayType, type='l', layout=c(1,2),
       xlab = '5-Min Interval',
       ylab = 'No. of Steps Taken')
 
````


